---
title: "Project2"
author: "Ben Spanswick"
date: "December 3, 2018"
output:
  html_document: default
  word_document: default
---


## Exectuive Summary

The goal of this project is to see if we can determine what causes attrition. We were provided a test set and a training set of data
for a number of employees. THe guidelines were to conduct the EDA, modify the data as necessary, and then build 3 models. 

The models are as follows:

1 - Simple linear model

2 - Model Studied in class (in this scenario I used KNN)

3 - Custom model. (In this scenario I built a ML pipeline in python, please see the git)

With the custom model, we were able to achieve a 88% accuracy rating on the test set. (.822 AUC), a result which I am very happy with.

For additonal context, please see the git readme file. 


---------------



Lets grab the data, and import some pakcages to start:


```{r}

data = read.csv("CaseStudy2-data.csv", header = TRUE)
pred = read.csv("CaseStudy2Validation.csv", header = TRUE)

data_original <- data
pred_original <- pred
data2 <-data

#install.packages("Amelia")
#install.packages('lemon')
#install.packages("mltools")
#install.packages("xlsx")


library(mltools) #mltools for one hot encoding
library(data.table)
library(Amelia)
library(ggplot2) 
library(tidyr)
library(purrr)
library(lemon) #prints prettyer tables
library(qtlcharts) #interactive plots - corr plot in this case
library(corrplot) #package corrplot
library(caret)
library(xlsx)
library(class)


knit_print.data.frame <- lemon_print

```


EDA: 

First lets look at the dimensions and check for missing varaibles. 


```{r}

#creating a simple miss map. 

missmap(data,
        main = "Missing Attrition Data",
        x.cex = 0.55,
        y.labels  = NULL,
        y.at = NULL
  )


```


No missing observations. Great!

And now for data structrue: 1170 Observations and 37 variables. 


```{r}

#Check Data Structure. 

str(data)

```


```{r}

#Here we need to convert some of the survey results into fators.
#I'm not actually sure if we need to do this, but I did, addtitional research will but this is how I did it this time around. 

data$BusinessTravel <- as.factor(data$BusinessTravel)
data$Gender <- as.factor(data$Gender)
data$MaritalStatus <- as.factor(data$MaritalStatus)
data$OverTime <- as.factor(data$OverTime)
data$WorkLifeBalance <- as.factor(data$WorkLifeBalance)
data$EnvironmentSatisfaction <- as.factor(data$EnvironmentSatisfaction)
data$JobInvolvement <- as.factor(data$JobInvolvement)
data$JobSatisfaction <- as.factor(data$JobSatisfaction)
data$PerformanceRating <- as.factor(data$PerformanceRating)
data$RelationshipSatisfaction <- as.factor(data$RelationshipSatisfaction)
data$JobLevel <- as.factor(data$JobLevel)
```




```{r}

# Checking levels of Job Role.


levels(data$JobRole)
```


```{r}

#Checking levels of Educational field.

levels(data$EducationField)

```

```{r}

#Checking levels of Department. 

levels(data$Department)
```



```{r}

#Transforming some of the variables to 1/0 or to other dummy variables. 

data$Attrition <- ifelse(data$Attrition == "Yes", 1, 0)
data$Gender <- ifelse(data$Gender == "Male", 1, 0)
data$OverTime <- ifelse(data$Gender == "Yes", 1, 0)
data$BusinessTravel <- ifelse(data$BusinessTravel == "Non-Travel", -1, ifelse(data$BusinessTravel == "Travel_Frequently", 1, 0))
data$MaritalStatus <- ifelse(data$MaritalStatus == "Divorced", -1, ifelse(data$MaritalStatus == "Married", 1, 0))

```
```{r}
#Quick look at the top of the data. 

head(data, 3)

```


```{r}


#For the KNN model, I actually had to go back and transform some of the Prediction data so that it was the same as the test dataset.Thats what this section is doing. 


Pred_NF <- pred

Pred_NF$BusinessTravel <- as.factor(Pred_NF$BusinessTravel)
Pred_NF$Gender <- as.factor(Pred_NF$Gender)
Pred_NF$MaritalStatus <- as.factor(Pred_NF$MaritalStatus)
Pred_NF$OverTime <- as.factor(Pred_NF$OverTime)
Pred_NF$WorkLifeBalance <- as.factor(Pred_NF$WorkLifeBalance)
Pred_NF$EnvironmentSatisfaction <- as.factor(Pred_NF$EnvironmentSatisfaction)
Pred_NF$JobInvolvement <- as.factor(Pred_NF$JobInvolvement)
Pred_NF$JobSatisfaction <- as.factor(Pred_NF$JobSatisfaction)
Pred_NF$PerformanceRating <- as.factor(Pred_NF$PerformanceRating)
Pred_NF$RelationshipSatisfaction <- as.factor(Pred_NF$RelationshipSatisfaction)
Pred_NF$JobLevel <- as.factor(Pred_NF$JobLevel)


Pred_NF$Attrition <- ifelse(Pred_NF$Attrition == "Yes", 1, 0)
Pred_NF$Gender <- ifelse(Pred_NF$Gender == "Male", 1, 0)
Pred_NF$OverTime <- ifelse(Pred_NF$Gender == "Yes", 1, 0)
Pred_NF$BusinessTravel <- ifelse(Pred_NF$BusinessTravel == "Non-Travel", -1, ifelse(Pred_NF$BusinessTravel == "Travel_Frequently", 1, 0))
Pred_NF$MaritalStatus <- ifelse(Pred_NF$MaritalStatus == "Divorced", -1, ifelse(Pred_NF$MaritalStatus == "Married", 1, 0))


```



```{r}

#Seperating out some numeric dataframes and some factor dataframes. 


NF <- data[, sapply(data, is.numeric)]
CF <- data[, sapply(data, is.factor)]

Pred_NF <- Pred_NF[, sapply(Pred_NF, is.numeric)]

```

```{r}

#Pariwise distribution plots. 


NF %>%
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
  facet_wrap(~ key, scales = "free") +   # In separate panels
  geom_histogram()


```

```{r}

#Corr plots. 

cors <- cor(NF)
corrplot(cors, method = "circle") #standard ugly correlation matrix

```



```{r}

#Now some interactive corr plots. 

NF$ID <-NULL
iplotCorr(NF) #create interactive qtlchart for correlation

Pred_NF$ID <-NULL

```

```{r}

#Dropping some columns we deem to be heavily corelated. 

drops <- c("YearsWithCurrManager","YearsSinceLastPromotion", "YearsInCurrentRole", "TotalWorkingYears", "JobLevel", "PercentSalaryHike", "EmployeeNumber", "EmployeeCount", "Rand", "Over18", "StandardHours", "Row.names", "Rand", "OVer18_Y", "EmployeeCount", "Department")
NF <- NF[ , !(names(NF) %in% drops)]
CF <- CF[ , !(names(CF) %in% drops)]
Pred_NF <- Pred_NF[ , !(names(Pred_NF) %in% drops)]
```

```{r}

#Fresh interactive corr plot. 

iplotCorr(NF)

```

```{r}

#Ugly corr plot again.

cors <- cor(NF)
corrplot(cors, method = "circle")
```

```{r}

#Lets check some VIFs scores. 

library(car)

vif(lm(Attrition ~JobRole + EducationField + Department, data=data))

```

```{r}

#Some one hot encoding which I ended up not using 

CF_DT <- as.data.table(CF)
CF1h <- as.data.frame(one_hot(CF_DT))

data_1h <- merge(NF, CF1h, by.x = 0, by.y = 0, all.x = TRUE, all.y = TRUE)

data_new <- merge(NF, CF, by.x = 0, by.y = 0, all.x = TRUE, all.y = TRUE)

head(data_1h, 3)


```

```{r}

#Creating some backup data so that I dont accidentally delete something. 

data_new <- data_new[ , !(names(data_new) %in% drops)]

rownames(data_new) <- data_new$ID

data_new <- data_new[,-1]


data3 <- data_new

str(data_new)

```

```{r}

#Again, this is the one hot encoded data, didnt bother with it, but leaving it here for reference for my future work. 


data_1h <- data_1h[ , !(names(data_1h) %in% drops)] #Now we have our final working data frame!

```




```{r}

#Checking the variable importance using GLM

train_control <- trainControl(method="cv", number=5)

data3$Attrition = as.factor(data3$Attrition)

glmnetTune <- train(Attrition~., data=data3, trControl=train_control, method="glmnet")

## absolute values in current version
varImp(glmnetTune, scale = FALSE, competes = FALSE)

## compared to coefficients with the sign values
multiCoefs = lapply(coef(glmnetTune$finalModel, s=glmnetTune$finalModel$tuneValue$lambda),   
        FUN=function (x) as.data.frame(t(as.matrix(x))))


```

```{r}

#FInally building the models, both KNN and Simple linear model. 


simple_mod <- glm(Attrition ~ I(OverTime == 'Yes') + I(JobRole == 'Research Director') + I(JobRole == 'Sales Representative'), data=data3, family='binomial')

#simple_mod2 <- glm(Attrition ~ OverTime + JobRole + JobInvolvement, data=data3, family='binomial')

#simple_mod3 <- glm(Attrition ~ I(OverTime == 'Yes') + I(JobRole == 'Research Director') + JobInvolvement, data=data2, family='binomial')


knn.train<- NF[,]
knn.train.target<- NF[,2]
knn.test<- Pred_NF[,]
knn.test.target<- Pred_NF[,2]


knn.10 <- knn(train=knn.train, test=knn.test, cl = knn.train.target, k=10)


pred$JobInvolvement <- as.factor(pred$JobInvolvement)


p <- predict(simple_mod, pred, type="response")

attrited  <-ifelse( p > .51, "yes", "no")

knn.attrited  <-ifelse( p == 1, "yes", "no")


output <- cbind(pred, attrited)

knn.output <- cbind(Pred_NF, knn.test.target)



```

```{r}

#Printing out the results. 

write.xlsx(output, "simple_model_output.xlsx")

write.xlsx(knn.output, "knn_output.xlsx")


```





